---
title: "igrf use"
author: "Koen Hufkens"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{igrf use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# load the library
library(igrf)
library(ggplot2)
library(metR)
```

### Single coordinates

IGRF values can be calculated for a single point using the below call. This will generate a data frame with model values. The routine is relatively fast so looping over a time vector will generate time series fairly quickly. To calculate grids a simple function is provided (see below).

```{r}
df <- igrf(
  field = "main",
  year = 2000,
  altitude = 2,
  latitude = 0,
  longitude = 0
)

print(df)
```

### Regular grid

You may generate global maps of the IGRF using the `igrf_grid()` function, setting similar parameters as above while specifying a resolution as decimal degrees of the global grid. Keep in mind that values under 1 (fractions) will take an increasing amount of time and space to store the data.

```{r}
grid <- igrf::igrf_grid(
  field = "main",
  year = 2000,
  altitude = 2,
  resolution = 5
  )

print(head(grid))
```

```{r  message=FALSE, results='hide'}
library(ggplot2)
library(metR)
library(rnaturalearth)

coast <- ne_coastline(returnclass = "sf")

ggplot(grid) +
  geom_contour_fill(
    aes(
      lon,
      lat,
      z = D
    ),
    breaks = MakeBreaks(10)
  ) +
  geom_sf(
    data = coast) +
  scale_fill_divergent(
    name = "declination"
  ) +
  labs(
    x = "",
    y = "",
    title = "IGRF magnetic declination"
  ) +
  theme_bw() +
  theme(
    plot.background = element_rect(fill = "#ffffff", color = NA),
    panel.background = element_rect(fill = "#ffffff", color = NA)
  )
```
